### 2019-12-10 analysis/R_scripts/select_covariates.Rmd
added script to select QRF covariates and make figures for manuscript.

---
title: "QRF Covariate Selection"
author: "Kevin See"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

# load packages for analysis
library(QRFcapacity)
library(maptools)
library(tidyverse)
library(janitor)
library(magrittr)
library(minerva)
library(sf)
library(quantregForest)
library(survey)
library(MuMIn)

# set default theme for ggplot
theme_set(theme_bw())
# setwd('analysis')
```

We loaded the data that had been prepped and saved as part of the `QRFcapacity` package. We used the data from all CHaMP sites, 2011-2017.

```{r load-data}
# determine which set of fish/habitat data to use
data("fh_sum_champ_2017")
fish_hab = fh_sum_champ_2017 %>%
  mutate_at(vars(Watershed, Year),
            list(as.factor))

# and the appropriate habitat dictionrary to go with it
data("hab_dict_2017")
hab_dict = hab_dict_2017

# all the related habitat data
data("champ_site_2011_17")
hab_data = champ_site_2011_17

data("champ_site_2011_17_avg")
hab_avg = champ_site_2011_17_avg

# add temperature metrics
data("champ_temps")

# add average aug temperature to average CHaMP metrics
hab_avg %<>%
  left_join(champ_temps %>%
              as_tibble() %>%
              select(Site, avg_aug_temp = S2_02_11) %>%
              distinct())

# add specific years' temperature, and average for each site
hab_data %<>%
  left_join(hab_data %>%
              select(VisitID, Year = VisitYear) %>%
              distinct() %>%
              left_join(champ_temps %>%
                          as_tibble() %>%
                          select(VisitID, avg_aug_temp = S2_02_11))) %>%
  left_join(hab_data %>%
              select(VisitID, Year = VisitYear) %>%
              distinct() %>%
              left_join(champ_temps %>%
                          as_tibble() %>%
                          select(Site:VisitID, S1_93_11:S36_2015) %>%
                          gather(scenario, aug_temp, S1_93_11:S36_2015) %>%
                          mutate(Year = str_sub(scenario, -4)) %>%
                          mutate_at(vars(Year),
                                    list(as.numeric)) %>%
                          filter(!is.na(Year)) %>%
                          select(Site:VisitID, Year, aug_temp))) %>%
  select(-Year)

```

All sites fell within the range of steelhead, but many of them were outside the range of spring/summer Chinook salmon. However, defining the range of sp/su Chinook is tricky, and it is unclear if the reported data for Chinook salmon includes only sites considered with the range of Chinook, or if we should exclude data from some sites based on our best understanding of the species' range. Currently, we are excluding sites outside the Chinook domain. We created our own list of sites in the John Day that were within a Chinook domain, based on which sites Chinook were found at some point (and downstream of those).

```{r chnk-domain-filter}
data("chnk_domain")

# which sites were sampled for Chinook? 
chnk_samps = fish_hab %>%
  filter(Species == 'Chinook') %>%
  select(Site:Lon, N) %>%
  distinct() %>%
  st_as_sf(coords = c('Lon', 'Lat'),
           crs = 4326) %>%
  st_transform(st_crs(chnk_domain))

# set snap distance (in meters)
st_crs(chnk_samps)
snap_dist = 1000

# which of those sites are in Chinook domain?
chnk_sites = chnk_samps %>%
  as_Spatial() %>%
  maptools::snapPointsToLines(chnk_domain %>%
                                mutate(id = 1:n()) %>%
                                select(id, MPG) %>%
                                as_Spatial(),
                              maxDist = snap_dist,
                              withAttrs = T,
                              idField = 'id') %>%
  as('sf') %>%
  select(-nearest_line_id, -snap_dist) %>%
  # include some sites in the John Day where Chinook were found (or seemed to be close to sites where Chinook were found)
  rbind(st_read('../data/derived_data/Chnk_JohnDay_TrueObs.shp',
                quiet = T) %>%
          st_transform(st_crs(chnk_domain)) %>%
          select(-in_range))

fish_hab %<>%
  filter(Species == 'Steelhead' |
           (Species == 'Chinook' & Site %in% chnk_sites$Site))
```

Next we generated MINE stats for all possible habitat metrics, matching them up against fish densities.

```{r MINE_stats}

# # what are some possible habitat covariates?
poss_hab_mets = hab_dict %>%
  filter(MetricCategory != 'Categorical') %>%
  filter(ShortName %in% names(fish_hab)) %>%
  pull(ShortName) %>%
  unique()

poss_hab_mets = c(poss_hab_mets, 'aug_temp', 'avg_aug_temp') %>%
  unique()

mine_res = fish_hab %>%
  split(list(.$Species)) %>%
  map_df(.id = 'Species',
         .f = function(x) {
           x %>%
             mutate(fish_dens = log(fish_dens + 0.005)) %>%
             estimate_MIC(covars = poss_hab_mets,
                          response = 'fish_dens')
         }) %>%
  left_join(hab_dict %>%
              filter(MetricGroupName == 'Visit Metric') %>%
              select(Metric = ShortName,
                     MetricCategory,
                     Name),
            by = 'Metric') %>%
  mutate_at(vars(MetricCategory),
            list(fct_explicit_na),
            na_level = 'Other') %>%
  mutate_at(vars(Name),
            list(as.character)) %>%
  mutate(Name = if_else(is.na(Name),
                        as.character(Metric),
                        Name)) %>%
  # put the metric names in descending order by MIC
  mutate_at(vars(Metric, Name),
            list(~ fct_reorder(., .x = MIC))) %>%
  select(Species, MetricCategory, Metric, everything()) %>%
  arrange(Species, MetricCategory, desc(MIC))

mine_plot_df = mine_res #%>%
  # filter out some metrics with too many NAs or 0s
  # filter((perc_NA < 0.2 & non_0 > 100) | MetricCategory == 'Temperature') #%>%
  # # filter out metrics with very low variance
  # filter(var < 0.1) %>% select(1:11)
  # # filter out area and volume metrics
  # filter(!grepl('Area', Metric),
  #        !grepl('Vol', Metric),
  #        Metric != 'Lgth_Wet')
```