### 2019-12-10 analysis/R_scripts/fit_qrf_model.Rmd
renamed script to fit QRF models.

---
title: "QRF Covariate Selection"
author: "Kevin See"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

# load packages for analysis
library(QRFcapacity)
library(maptools)
library(tidyverse)
library(janitor)
library(magrittr)
library(minerva)
library(sf)
library(quantregForest)
library(survey)
library(MuMIn)

# set default theme for ggplot
theme_set(theme_bw())
# setwd('analysis/R_scripts')
```

# Select QRF Habitat Covariates

We loaded the data that had been prepped and saved as part of the `QRFcapacity` package. We used the data from all CHaMP sites, 2011-2017.

```{r load-data}
# determine which set of fish/habitat data to use
data("fh_sum_champ_2017")
fish_hab = fh_sum_champ_2017 %>%
  filter(Species == 'Chinook') %>%
  mutate_at(vars(Watershed, Year),
            list(as.factor))

# and the appropriate habitat dictionrary to go with it
data("hab_dict_2017")
hab_dict = hab_dict_2017

# all the related habitat data
data("champ_site_2011_17")
hab_data = champ_site_2011_17 %>%
  filter(!Watershed %in% c('Big-Navarro-Garcia (CA)',
                           'CHaMP Training',
                           'Region 17',
                           'Walla Walla',
                           'Umatilla'))

data("champ_site_2011_17_avg")
hab_avg = champ_site_2011_17_avg %>%
  filter(!Watershed %in% c('Big-Navarro-Garcia (CA)',
                           'CHaMP Training',
                           'Region 17',
                           'Walla Walla',
                           'Umatilla'))

# add temperature metrics
data("champ_temps")

# add average aug temperature to average CHaMP metrics
hab_avg %<>%
  left_join(champ_temps %>%
              as_tibble() %>%
              select(Site, avg_aug_temp = S2_02_11) %>%
              distinct())

# add specific years' temperature, and average for each site
hab_data %<>%
  left_join(hab_data %>%
              select(VisitID, Year = VisitYear) %>%
              distinct() %>%
              left_join(champ_temps %>%
                          as_tibble() %>%
                          select(VisitID, avg_aug_temp = S2_02_11))) %>%
  left_join(hab_data %>%
              select(VisitID, Year = VisitYear) %>%
              distinct() %>%
              left_join(champ_temps %>%
                          as_tibble() %>%
                          select(Site:VisitID, S1_93_11:S36_2015) %>%
                          gather(scenario, aug_temp, S1_93_11:S36_2015) %>%
                          mutate(Year = str_sub(scenario, -4)) %>%
                          mutate_at(vars(Year),
                                    list(as.numeric)) %>%
                          filter(!is.na(Year)) %>%
                          select(Site:VisitID, Year, aug_temp))) %>%
  select(-Year)

```

All sites fell within the range of steelhead, but many of them were outside the range of spring/summer Chinook salmon. However, defining the range of sp/su Chinook is tricky, and it is unclear if the reported data for Chinook salmon includes only sites considered with the range of Chinook, or if we should exclude data from some sites based on our best understanding of the species' range. Currently, we are excluding sites outside the Chinook domain. We created our own list of sites in the John Day that were within a Chinook domain, based on which sites Chinook were found at some point (and downstream of those).

```{r chnk-domain-filter}
data("chnk_domain")

# which sites were sampled for Chinook? 
chnk_samps = fish_hab %>%
  filter(Species == 'Chinook') %>%
  select(Site:Lon, N) %>%
  distinct() %>%
  st_as_sf(coords = c('Lon', 'Lat'),
           crs = 4326) %>%
  st_transform(st_crs(chnk_domain))

# set snap distance (in meters)
st_crs(chnk_samps)
snap_dist = 1000

# which of those sites are in Chinook domain?
chnk_sites = chnk_samps %>%
  as_Spatial() %>%
  maptools::snapPointsToLines(chnk_domain %>%
                                mutate(id = 1:n()) %>%
                                select(id, MPG) %>%
                                as_Spatial(),
                              maxDist = snap_dist,
                              withAttrs = T,
                              idField = 'id') %>%
  as('sf') %>%
  select(-nearest_line_id, -snap_dist) %>%
  # include some sites in the John Day where Chinook were found (or seemed to be close to sites where Chinook were found)
  rbind(st_read('../data/derived_data/Chnk_JohnDay_TrueObs.shp',
                quiet = T) %>%
          st_transform(st_crs(chnk_domain)) %>%
          select(-in_range))

fish_hab %<>%
  filter(Site %in% chnk_sites$Site)
  # filter(Species == 'Steelhead' |
  #          (Species == 'Chinook' & Site %in% chnk_sites$Site))

```

Next we generated MINE stats for all possible habitat metrics, matching them up against the log of fish densities (plus a small value to deal with 0s).

```{r MINE_stats}

# # what are some possible habitat covariates?
poss_hab_mets = hab_dict %>%
  filter(MetricCategory != 'Categorical') %>%
  filter(ShortName %in% names(fish_hab)) %>%
  pull(ShortName) %>%
  unique()

# poss_hab_mets = c(poss_hab_mets, 'aug_temp', 'avg_aug_temp') %>%
#   unique()

mine_res = fish_hab %>%
  mutate_at(vars(starts_with('LWVol')),
            list(~ . / Lgth_Wet)) %>%
  mutate(fish_dens = log(fish_dens + 0.005)) %>%
  estimate_MIC(covars = poss_hab_mets,
               response = 'fish_dens') %>%
  mutate(Species = 'Chinook') %>%
  left_join(hab_dict %>%
              filter(MetricGroupName == 'Visit Metric') %>%
              select(Metric = ShortName,
                     MetricCategory,
                     Name),
            by = 'Metric') %>%
  mutate_at(vars(MetricCategory),
            list(fct_explicit_na),
            na_level = 'Other') %>%
  mutate_at(vars(Name),
            list(as.character)) %>%
  mutate(Name = if_else(is.na(Name),
                        as.character(Metric),
                        Name)) %>%
  # put the metric names in descending order by MIC
  mutate_at(vars(Metric, Name),
            list(~ fct_reorder(., .x = MIC))) %>%
  select(Species, MetricCategory, Metric, everything()) %>%
  arrange(Species, MetricCategory, desc(MIC))

```

Plot the MINE results and save figures for manuscript.

```{r MINE-figures}
mine_plot_df = mine_res %>%
  # filter out some metrics with too many NAs or 0s
  filter((perc_NA < 0.2 & non_0 > 100) | MetricCategory == 'Temperature') %>%
  # filter out metrics with very low variance
  # filter(var < 0.1) %>%
  # filter(obsCV < 0.1)
  # janitor::tabyl(MetricCategory)
  # select(1:11)
  # filter out area and volume metrics
  filter(!grepl('Area$', Metric),
         !grepl('Vol$', Metric),
         !Metric %in% c('Lgth_Wet', 
                        'Lgth_BfChnl',
                        'Lgth_WetChnl',
                        'Area_Wet', 
                        'Area_Bf', 
                        'WetVol', 
                        'BfVol'))

mine_p = mine_plot_df %>%
  ggplot(aes(x = Name,
             y = MIC)) +
  geom_col(position = position_dodge(1),
           fill = 'gray45') +
  coord_flip() +
  facet_wrap(~ MetricCategory,
             scales = 'free_y',
             ncol = 3) +
  scale_fill_brewer(palette = 'Set1',
                    guide = guide_legend(nrow = 1)) +
  theme(legend.position = 'bottom',
        axis.text = element_text(size = 5)) +
  labs(y = 'Habitat Covariate')

mine_p2 = mine_plot_df %>%
  ggplot(aes(x = Name,
             y = MIC,
             fill = MetricCategory)) +
  geom_col() +
  coord_flip() +
  scale_fill_brewer(palette = 'Set1',
                    guide = guide_legend(nrow = 1)) +
  theme(legend.position = 'bottom',
        axis.text = element_text(size = 5)) +
  labs(y = 'Habitat Covariate')

```

```{r, eval = F}
library(corrr)
#----------------------------------------------
# Look at correlations between habitat metrics
#----------------------------------------------
# top metrics
sel_mets = poss_hab_mets
sel_mets = mine_plot_df %>%
  group_by(MetricCategory) %>%
  slice(1:5) %>%
  ungroup() %>%
  pull(Metric) %>%
  unique() %>%
  as.character()

corr_mat = hab_avg %>%
  select(one_of(sel_mets)) %>%
  corrr::correlate()

corr_mat %>%
  corrr::rearrange(absolute = F) %>%
  corrr::shave(upper = T) %>% 
  corrr::stretch() %>%
  filter(!is.na(r)) %>%
  filter(abs(r) > 0.5) %>%
  kable()

corr_p1 = corr_mat %>%
  # rearrange(absolute = F) %>%
  shave(upper = T) %>% 
  rplot(legend = T,
        print_cor = T)

corr_p2 = network_plot(corr_mat)
```

```{r correlation-plots}
#----------------------------------------------
# Look at correlations between habitat metrics
#----------------------------------------------

library(ggcorrplot)

corr_list = hab_dict %>%
  filter(MetricCategory != 'Categorical') %>%
  split(list(.$MetricCategory)) %>%
  map(.f = function(x) {
    catg_mets = x %>%
      pull(ShortName)
    hab_data %>%
      select(one_of(catg_mets)) %>%
      cor(use = "pairwise.complete.obs")
      # corrr::correlate(use = "pairwise.complete.obs",
      #                  method = "pearson",
      #                  quiet = T)
  })

# make correlation plots by metric category
corr_p_list = corr_list %>%
  map(.f = ggcorrplot,
      method = "circle",
      show.diag = F,
      type = 'lower')

# add a title to each plot
for(i in 1:length(corr_p_list)) {
  corr_p_list[[i]] = corr_p_list[[i]] +
    labs(title = names(corr_list)[i])
}

# save as a PDF
pdf("../figures/correlation_plots_all.pdf",
    width = 8,
    height = 8)
for(i in 1:length(corr_p_list)) {
  print(corr_p_list[[i]])
}
dev.off()

```

Based on the MIC results, and trying to stay away from pairs of metrics that were too closely correlated, we came up with the following list of QRF covariates.

```{r habitat-covariates}
sel_hab_mets = tibble(Species = c('Chinook'),
                      Metric = c('UcutArea_Pct',
                                 'FishCovNone',
                                 'SubEstGrvl',
                                 'FstTurb_Freq',
                                 'FstNT_Freq',
                                 'CU_Freq',
                                 'SlowWater_Pct',
                                 'NatPrin1',
                                 'DistPrin1',
                                 'avg_aug_temp',
                                 # 'Sin_CL',
                                 'Sin',
                                 'WetWdth_CV',
                                 'WetBraid',
                                 'WetSC_Pct',
                                 'Q',
                                 'WetWdth_Int',
                                 'LWFreq_Wet',
                                 'LWVol_WetFstTurb'))

# sum(!sel_hab_mets$Metric %in% names(fish_hab))
```

```{r}
sel_hab_mets %>%
  select(ShortName = Metric) %>%
  distinct() %>%
  left_join(hab_dict) %>%
  select(Catg = MetricCategory, 
         Metric = Name, 
         Description = DescriptiveText) %>%
  kable()
```

# Fit QRF Model

```{r}
#-----------------------------------------------------------------
# Fit QRF model
#-----------------------------------------------------------------
covars = sel_hab_mets$Metric
# impute missing data in fish / habitat dataset
set.seed(7)
qrf_data = fish_hab %>%
  impute_missing_data(covars = covars,
                      impute_vars = c('Watershed', 'Elev_M', 'Sin', 'Year', 'CUMDRAINAG'),
                      method = 'missForest') %>%
  select(Site, Watershed, Year, LON_DD, LAT_DD, fish_dens, VisitID, one_of(covars))

# fit the QRF model
# set the density offset (to accommodate 0z)
dens_offset = 0.005

# fit random forest models
set.seed(4)
qrf_mod = quantregForest(x = qrf_data %>%
                               select(one_of(covars)) %>%
                               as.matrix,
                             y = qrf_data %>%
                               mutate_at(vars(fish_dens),
                                         list(~ log(. + dens_offset))) %>%
                               select(fish_dens) %>%
                               as.matrix(),
                             keep.inbag = T,
                             ntree = 1000)

# save some results
use_data(fish_hab, 
         hab_dict,
         hab_data,
         hab_avg,
         sel_hab_mets,
         qrf_data,
         dens_offset,
         qrf_mod,
         version = 2,
         overwrite = T)

```

## Predict at all CHaMP sites

```{r champ_predictions}
# what quantile is a proxy for capacity?
pred_quant = 0.9

covars = unique(sel_hab_mets$Metric)

# impute missing values in habitat data
hab_impute = hab_avg %>%
  mutate_at(vars(Watershed, Channel_Type),
            list(fct_drop)) %>%
  impute_missing_data(data = .,
                      covars = covars,
                      impute_vars = c('Watershed', 
                                      'Elev_M', 
                                      'Channel_Type', 
                                      'CUMDRAINAG'),
                      method = 'missForest') %>%
  select(Site, Watershed, LON_DD, LAT_DD, VisitYear, Lgth_Wet, Area_Wet, one_of(covars))

# predict at all CHaMP sites
pred_hab_sites = hab_impute %>%
  mutate(chnk_per_m = predict(qrf_mod,
                              newdata = select(., one_of(covars)),
                              what = pred_quant),
         chnk_per_m = exp(chnk_per_m) - dens_offset,
         chnk_per_m2 = chnk_per_m * Lgth_Wet / Area_Wet)

# filter out sites outside the Chinook domain
data("chnk_domain")
snap_dist = 1000

pred_hab_sites_chnk = pred_hab_sites %>%
  filter(!is.na(LON_DD)) %>%
  st_as_sf(coords = c('LON_DD', 'LAT_DD'),
           crs = 4326) %>%
  st_transform(crs = st_crs(chnk_domain)) %>%
  as_Spatial() %>%
  maptools::snapPointsToLines(chnk_domain %>%
                                mutate(id = 1:n()) %>%
                                select(id, MPG) %>%
                                as_Spatial(),
                              maxDist = snap_dist,
                              withAttrs = T,
                              idField = 'id') %>%
  as('sf') %>%
  as_tibble() %>%
  select(-nearest_line_id, -snap_dist, -geometry) %>%
  # add sites in the John Day
  bind_rows(st_read('../data/derived_data/Chnk_JohnDay_TrueObs.shp',
                    quiet = T) %>%
              as_tibble() %>%
              select(Site) %>%
              inner_join(pred_hab_sites))

```

# Extrapolation