### 2020-01-28 analysis/R_scripts/compile_wtsd_bndry.Rmd
created R notebook to read in CHaMP watershed boundaries and save them as .rds object.

---
title: "Create CHaMP map"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

# load packages for analysis
library(tidyverse)
library(janitor)
library(magrittr)
library(sf)
library(ggrepel)

# set default theme for ggplot
theme_set(theme_bw())
# setwd('analysis/R_scripts')
```

Start by looking at which watersheds we need polygon shapes for.

```{r}
fish_hab = read_rds("../data/derived_data/fish_hab.rds")
model_svy_df = read_rds("../data/derived_data/model_svy_df.rds")
```

Watersheds with paired fish-habitat data:

```{r}
tabyl(fish_hab,
      Watershed) %>%
  adorn_pct_formatting()
```

Watersheds with CHaMP habitat data

```{r}
model_svy_df$data[[1]] %>%
  tabyl(Watershed) %>%
  adorn_pct_formatting()
```

Read in some HUC boundaries.

```{r}
huc3 = read_sf('~/OneDrive - Merck Sharp & Dohme, Corp/Data/WatershedBoundaries/WBDHU6.shp') %>%
  select(STATES:NAME) %>%
  st_transform(crs = 5070)
huc4 = read_sf('~/OneDrive - Merck Sharp & Dohme, Corp/Data/WatershedBoundaries/WBDHU8.shp') %>%
  select(STATES:NAME) %>%
  st_transform(st_crs(huc3))
huc5 = read_sf('~/OneDrive - Merck Sharp & Dohme, Corp/Data/WatershedBoundaries/WBDHU10.shp') %>%
  select(STATES:NAME) %>%
  st_transform(st_crs(huc3))

# filter(huc4,
#        grepl('Methow', NAME)) %>%
#   select(NAME) %>% plot()

champ_wtsds = huc3 %>%
  filter(NAME %in% c('John Day')) %>%
  select(NAME) %>%
  rbind(huc4 %>%
          filter(NAME %in% c('Lemhi',
                             'Methow',
                             'Wenatchee',
                             'South Fork Salmon',
                             'Upper Grande Ronde')) %>%
          select(NAME)) %>%
  rbind(huc5 %>%
          filter(NAME %in% c('Minam River',
                             'Secesh River',
                             'Yankee Fork')) %>%
          select(NAME) %>%
          mutate(NAME = str_remove(NAME, ' River$'))) %>%
  rbind(huc5 %>%
          filter(NAME %in% c('Entiat River',
                             'Mad River')) %>%
          st_union() %>%
          st_sf() %>%
          mutate(NAME = 'Entiat')) %>%
  rbind(huc5 %>%
          filter(grepl('Tucannon', NAME)) %>%
          st_union() %>%
          st_sf() %>%
          mutate(NAME = 'Tucannon')) %>%
  arrange(NAME) %>%
  mutate(id = 1:n(),
         hab_data = T,
         fish_data = if_else(NAME %in% c('Secesh', as.character(unique(fish_hab$Watershed))),
                             T, F),
         fish_data = if_else(NAME == 'South Fork Salmon', F, fish_data))

```

Test a plot

```{r}
ggplot(data = champ_wtsds) +
  geom_sf() +
  geom_sf(data = champ_wtsds %>%
            filter(fish_data),
          fill = 'black') +
  geom_sf_label(aes(label = id),
                position = position_dodge(width = 200)) +
                # position = position_nudge(x = 5000,
                #                           y = -5000)) +
  theme(axis.title = element_blank())

```

Looks good. Let's save the polygon sf to use later.

```{r}
write_rds(champ_wtsds,
          path = '../data/derived_data/champ_wtds.rds')
```