### 2020-06-09 analysis/paper/appendix_s1.Rmd
changed correlations to be done with Pearson correlation coefficients. Added some plots of selected metrics.

---
title: "Appendix S1"
subtitle: "Estimating Carrying Capacity for Chinook Parr using Quantile Random Forest Models"
author:
  - Kevin E. See:
      email: Kevin.See@biomark.com
      institute: [biomark]
      correspondence: true
  - Michael W. Ackerman:
      email: Mike.Ackerman@biomark.com
      institute: [biomark]
      correspondence: false
  - Richard A. Carmichael:
      email: Richie.Carmichael@biomark.com
      institute: [biomark]
      correspondence: false
  - Sarah L. Hoffmann:
      email: Sarah.Hoffmann@biomark.com
      institute: [biomark]
      correspondence: false
  - Chris Beasley:
      email: Chris.Beasley@biomark.com
      institute: [biomark]
      correspondence: false
institute:
  - biomark: Biomark, Inc. 705 South 8th St., Boise, Idaho, 83702, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      fig_height: 8.5
      fig_width: 6.5
      number_sections: FALSE
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::pdf_document2:
      fig_caption: yes
      fig_height: 8.5
      fig_width: 6.5
      toc: no
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
      includes:
        in_header: ../templates/header_appendixS.tex
bibliography:
- references.bib
- packages.bib
csl: "../templates/ecological-applications.csl" # Insert path for the bib-style
---

<!-- Find a way to add the journal name (Ecological Applications) to the top  -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/appendix/",
  dpi = 600,
  appendix = TRUE
)
```

```{r packages}
# setwd('analysis/paper')

library(tidyverse)
library(minerva)
library(QRFpaper)
library(kableExtra)

# set theme for ggplot2
theme_set(theme_bw())

# options for table formating with "kable()"
options(knitr.kable.NA = '-')
options(knitr.table.format = "pandoc")

```

```{r load-data}
fish_hab = read_rds("../data/derived_data/fish_hab.rds")
hab_dict = read_rds("../data/derived_data/hab_dict.rds")
sel_hab_mets = read_rds("../data/derived_data/sel_hab_mets.rds")
hab_data = read_rds("../data/derived_data/hab_data.rds")
```

```{r}
# what are some possible habitat covariates?
poss_hab_mets = tibble(ShortName = names(fish_hab)) %>%
  distinct() %>%
  left_join(hab_dict %>%
              filter(! MetricGroupName %in% c('Tier 1 Summary', 'Tier 2 Summary')) %>%
              select(ShortName, MetricGroupName, Name, MetricCategory) %>%
              distinct()) %>%
  filter(is.na(MetricCategory) | MetricCategory != 'Categorical') %>%
  filter((!is.na(Name) |
            ShortName %in% c('Elev_M',
                             'CUMDRAINAG',
                             "DpthThlwg_Avg",
                             "SCSm_Area",
                             "SCSm_Freq",
                             "SCSm_Ct",
                             "SCSm_Vol",
                             "RipCovUstoryNone",
                             "RipCovGrndNone",
                             "SC_Area",
                             "SC_Area_Pct",
                             "ChnlUnitTotal_Ct",
                             "Discharge_fish",
                             "Temp",
                             "PercentIceCover",
                             "LWCount",
                             "SubEstBdrk",
                             "Ucut_Length",
                             "FishCovAll",
                             "SubEstCandBldr",
                             "UcutLgth",
                             "LWcnt_Wet"))) %>%
  mutate(Name = if_else(is.na(Name),
                        ShortName,
                        Name)) %>%
  filter(!ShortName %in% c('Tier1', 'Tier2')) %>%
  mutate(MetricCategory = if_else(grepl('SC', ShortName),
                                  'SideChannel',
                                  MetricCategory),
         MetricCategory = if_else(grepl('Sub', ShortName),
                                  'Substrate',
                                  MetricCategory),
         MetricCategory = if_else(grepl('^Rip', ShortName),
                                  'Riparian',
                                  MetricCategory),
         MetricCategory = if_else(grepl('^FishCov', ShortName) |
                                    grepl('Ucut', ShortName) |
                                    ShortName %in% c('PercentIceCover'),
                                  'Cover',
                                  MetricCategory),
         MetricCategory = if_else(grepl('^LW', ShortName),
                                  'Wood',
                                  MetricCategory),
         MetricCategory = if_else(grepl('Discharge', ShortName) |
                                    ShortName %in% c("DpthThlwg_Avg",
                                                     "Dpth_Max",
                                                     'DpthThlwgExit',
                                                     'DpthResid',
                                                     'TotalVol',
                                                     'CUMDRAINAG'),
                                  'Size',
                                  MetricCategory),
         MetricCategory = if_else(ShortName %in% c('Elev_M', 'Temp'),
                                  'Temperature',
                                  MetricCategory),
         MetricCategory = if_else(ShortName %in% c('ChnlUnitTotal_Ct'),
                                  'ChannelUnit',
                                  MetricCategory)) %>%
  select(-MetricGroupName)

# # what columns in fish_hab are not in poss_hab_mets?
# names(fish_hab)[!names(fish_hab) %in% poss_hab_mets$ShortName]
```

```{r mine-stats}
#-----------------------------------------------------------------
# generate MINE statistics
#-----------------------------------------------------------------
mine_res = tibble(species = unique(fish_hab$Species),
                  fh_data = list(fish_hab)) %>%
  mutate(metrics = map(fh_data,
                       .f = function(x) {
                         poss_hab_mets %>%
                           filter(ShortName %in% names(x)) %>%
                           pull(ShortName)
                       })) %>%
  mutate(mine_res = map2(fh_data,
                         metrics,
                         .f = function(x, y) {
                           if(sum(x$fish_dens == 0) == 0) {
                             try(x %>%
                                   mutate(fish_dens = log(fish_dens)) %>%
                                   estimate_MIC(covars = y,
                                                response = 'fish_dens'))
                           } else {
                             try(x %>%
                                   mutate(fish_dens = log(fish_dens + 0.005)) %>%
                                   estimate_MIC(covars = y,
                                                response = 'fish_dens'))
                           }
                         })) %>%
  select(species, mine_res) %>%
  unnest(cols = mine_res) %>%
  left_join(poss_hab_mets,
            by = c('Metric' = 'ShortName')) %>%
  select(species, MetricCategory, Metric, Name, everything()) %>%
  # put the metric names in descending order by MIC
  mutate_at(vars(Metric, Name),
            list(~ fct_reorder(., .x = MIC))) %>%
  arrange(species, MetricCategory, desc(MIC)) %>%
      # filter out some metrics with too many NAs or 0s
      # filter((perc_NA < 0.5 & non_0 > 100)) %>%
      # filter out metrics with very low variance
      # filter(var < 0.1) %>%
      # filter(obsCV < 0.1)
      # filter out area and volume metrics
      filter(!grepl('Area$', Metric),
             !grepl('Vol$', Metric),
             !Metric %in% c('Lgth_Wet', 
                            'Lgth_BfChnl',
                            'Lgth_WetChnl',
                            'Area_Wet', 
                            'Area_Bf', 
                            'WetVol', 
                            'BfVol')) %>%
  mutate(selected = if_else(Metric %in% sel_hab_mets$Metric,
                            T, F))

```

```{r calc-correlations}
library(corrr)

hab_corr = tibble(hab_data = list(hab_data)) %>%
  mutate(metrics = map(hab_data,
                       .f = function(x) {
                         mets = poss_hab_mets %>%
                           filter(ShortName %in% names(x)) %>%
                           pull(ShortName)
                         x %>%
                           select(one_of(mets)) %>%
                           gather(Metric, value) %>%
                           group_by(Metric) %>%
                           summarise(n_tot = n(),
                                     n_NA = sum(is.na(value)),
                                     non_0 = sum(value != 0, na.rm = T)) %>%
                           mutate(perc_NA = n_NA / n_tot,
                                  perc_non_0 = non_0 / n_tot) %>%
                           filter(perc_NA < 0.5,
                                  non_0 > 100) %>%
                           filter(!grepl('Area$', Metric),
                                  !grepl('Vol$', Metric),
                                  !Metric %in% c('Lgth_Wet', 
                                                 'Lgth_BfChnl',
                                                 'Lgth_WetChnl',
                                                 'Area_Wet', 
                                                 'Area_Bf', 
                                                 'WetVol', 
                                                 'BfVol')) %>%
                           pull(Metric)
                       }),
         cor_mat = map2(hab_data,
                        metrics,
                        .f = function(x, y) {
                          x %>%
                            select(one_of(y)) %>%
                            cor(use = 'pairwise',
                                method = 'pearson')
                                # method = 'spearman')
                        }),
         corr_mat = map(cor_mat,
                        .f = as_cordf),
         catg_cor = map2(hab_data,
                         metrics,
                         .f = function(x, y) {
                           catg_list = try(poss_hab_mets %>%
                                             filter(ShortName %in% y) %>%
                                             split(list(.$MetricCategory)) %>%
                                             map(.f = function(z) {
                                               if(nrow(z) == 1) {
                                                 return(NULL)
                                               } else {
                                                 x %>%
                                                   select(one_of(z$ShortName)) %>%
                                                   cor(use = 'pairwise',
                                                       method = 'pearson')
                                                       # method = 'spearman')
                                               }
                                               }))
                           return(catg_list)
                                             }))

```

# Choosing Habitat Covariates

A key step in developing a QRF model to predict fish capacities was selecting the habitat covariates to include in the model. Random forest models naturally incorporate interactions between correlated covariates, which is essential since nearly all habitat variables are considered correlated to one degree or another, however, we aimed to avoid overly redundant variables (i.e., variables that measure similar aspects of the habitat). Further, including too many covariates can result in overfitting of the model (e.g., including as many covariates as data points).

To prevent overfitting, we pared down the more than 100 metrics generated by the CHaMP protocol describing the quantity and quality of fish habitat for each survey site. To assist in determining the habitat metrics to include in the QRF model, we used the Maximal Information-Based Nonparametric Exploration (MINE) class of statistics [@Reshef2011] to determine those habitat characteristics (covariates) most highly associated with observed parr densities. We calculated the maximal information coefficient (MIC), using the R package *minerva* [@R-minerva], to measure the strength of the linear or non-linear association between two variables [@Reshef2011]. The MIC value between each of the measured habitat characteristics and parr density was used to inform decisions on which habitat covariates to include in the QRF parr capacity model.

Habitat metrics were first grouped into broad categories that included channel unit, complexity, cover, disturbance, riparian, size, substrate, temperature, water quality, and woody debris. Habitat metrics measuring volume or area were scaled to the wetted area of each site. Within each category, metrics were ranked according to their MIC value (Table \@ref(tab:mic-tab) and Figure \@ref(fig:mine-facet-fig)). Our strategy was to select one or two variables with the highest MIC score within each category so that covariates describe different aspects of rearing habitat (e.g., substrate, temperature, etc.). Additionally, we attempted to avoid covariates that were highly correlated (**Figure ??**) while including covariates that can be directly influenced by restoration actions or have been shown to impact salmonid juvenile density.

<!-- talk about how we chose not to use categorical variables like channle type or valley class, because we expected other metrics to capture the difference between those categories -->

<!-- The following line inserts a page break  -->
\newpage

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

\newpage

# Tables

```{r mic-tab}
mine_tab_df = mine_res %>%
  group_by(MetricCategory) %>%
  slice(1:10) %>%
  select(Category = MetricCategory,
         Name,
         Abbrv = Metric,
         MIC,
         selected)

mine_tab_df %>%
  select(-selected) %>%
  kable(digits = 3,
        format = 'html',
        caption = 'MIC statistic for top metrics in each habitat category. Metrics selected for the QRF model are in bold.') %>%
  kable_styling(fixed_thead = T) %>%
  row_spec(row = which(mine_tab_df$selected),
           bold = T)

```

# Figures

```{r mine-facet-fig, fig.cap = 'Barplots of MIC statistics, faceted by habitat category.'}
ggplot(data = mine_res,
       aes(x = Name,
           y = MIC)) +
  geom_col(fill = 'blue') +
  coord_flip() +
  facet_wrap(~ MetricCategory,
             scales = 'free_y',
             ncol = 3) +
  theme(legend.position = 'bottom',
        axis.text = element_text(size = 5))
```

```{r mine-all-fig, fig.cap = 'Barplot of MIC statistics, colored by habitat category.'}
ggplot(data = mine_res,
       aes(x = Name,
           y = MIC,
           fill = MetricCategory)) +
  geom_col() +
  coord_flip() +
  # scale_fill_viridis_d() +
  scale_fill_brewer(palette = 'Set3',
                    guide = guide_legend(nrow = 2)) +
  theme(legend.position = 'bottom',
        axis.text = element_text(size = 6)) +
  labs(fill = 'Category')
```

```{r corr-fig, eval = F, results='hide', fig.width = 5, fig.height = 5, fig.show = "hold", fig.cap = "Correlation plots of metrics, facted by habitat category."}
hab_corr$catg_cor[[1]][!sapply(hab_corr$catg_cor[[1]], is.null)] %>%
  # map(.f = function(x) {
  #   rownames(x) <- tibble(Metric = rownames(x)) %>%
  #     left_join(hab_dict %>%
  #                 select(Metric = ShortName,
  #                        Name) %>%
  #                 distinct()) %>%
  #     pull(Name)
  #   
  #   # colnames(x) <- tibble(Metric = colnames(x)) %>%
  #   #   left_join(hab_dict %>%
  #   #               select(Metric = ShortName,
  #   #                      Name) %>%
  #   #               distinct()) %>%
  #   #   pull(Name)
  #   return(x)
  # }) %>%
    map(.f = corrplot::corrplot.mixed,
        upper = 'ellipse',
        tl.pos = 'lt',
        order = 'FPC')
```

```{r corr-fig2, results='hide', fig.width = 5, fig.height = 5, fig.show = "hold", fig.cap = "Correlation plots of metrics, facted by habitat category."}
cor_plot_df = tibble(Category = names(hab_corr$catg_cor[[1]][!sapply(hab_corr$catg_cor[[1]], is.null)]),
                     cor_mat = hab_corr$catg_cor[[1]][!sapply(hab_corr$catg_cor[[1]], is.null)]) %>%
  mutate(cor_plot = map2(cor_mat,
                         Category,
                         .f = function(x, y) {
                           ggcorrplot::ggcorrplot(x,
                                                  method = 'circle',
                                                  hc.order = T,
                                                  type = 'lower',
                                                  title = y,
                                                  colors = c('blue4', 'white', 'red3'))
                         }))

cor_plot_df$cor_plot
```

```{r selected-corr, fig.cap = 'Correlation plot of habitat metrics used in QRF model.'}
hab_data %>% 
  select(one_of(sel_hab_mets$Metric)) %>% 
  cor(use = 'pairwise',
      method = 'pearson') %>%
      # method = 'spearman') %>% 
  # as_cordf() %>%
  ggcorrplot::ggcorrplot(method = 'circle',
                         # hc.order = T,
                         type = 'lower',
                         # legend.title = "Correlation",
                         colors = c('blue4', 'white', 'red3'))
```

```{r high-cor-mets, fig.cap = 'Pairs plot of habitat metrics used in QRF model with a correlation coefficient greater than 0.5.'}
high_cor_mets = hab_data %>%
  select(one_of(sel_hab_mets$Metric)) %>%
  correlate(use = 'pairwise',
            method = 'pearson') %>%
            # method = 'spearman') %>%
  stretch(na.rm = T,
          remove.dups = T) %>%
  arrange(desc(abs(r))) %>%
  filter(abs(r) > 0.5) %>%
  select(x, y) %>%
  pivot_longer(cols = c(x, y)) %>%
  pull(value) %>%
  unique()

hab_data %>%
  select(one_of(high_cor_mets)) %>%
  GGally::ggpairs()
```