### 2021-03-09 README.Rmd
updated some references.
Added DOI info to the README page, now that the paper is published.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Estimating Carrying Capacity for Juvenile Salmon using Quantile Random Forest Models"
Authors <- "See, Kevin E., Michael W. Ackerman, Richard A. Carmichael, Sarah Hoffman, and Chris Beasley"
yr <- 2021
Journal <- "Ecosphere"
```

# QRFpaper

[![DOI](https://zenodo.org/badge/225946308.svg)](https://zenodo.org/badge/latestdoi/225946308)

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/KevinSee/QRFpaper/master?urlpath=rstudio)

This repository contains the data and code for our paper:

> `r Authors` (`r yr`). _`r Title`_. `r Journal` <http://doi.org/10.1002/ecs2.3404 >

### How to cite

Please cite this compendium as:

> `r Authors`, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.5281/zenodo.4300942>

### How to download or install

You can download the compendium as a zip from from this URL: <https://github.com/KevinSee/QRFpaper/archive/master.zip>

Or you can install this compendium as an R package, QRFpaper, from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
remotes::install_github("KevinSee/QRFpaper")
```

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.